Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-05-22 performance_comparison.tex
[MM] some captions

\begin{figure}
	\centering
	\includegraphics{Fig/gamma_sensitivity.pdf}
	\hfill
	\includegraphics{Fig/Pzero_sensitivity.pdf}
	\caption{}
	\label{fig:ol_sensitivities}
\end{figure}

\subsection{Robustified constraints comparison}
It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style. This analysis compares the nominal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC). All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of meeting both constraints.

Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the nominal trajectory, for the three previously mentioned cases. The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel). Except for the $u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal-whether it is positive, negative, or numerically zero. These sign indicators are represented as three distinct levels---high, zero, and low---corresponding to positive, zero, and negative values of the nominal signal, respectively. The levels are scaled appropriately in each plot to ensure clarity of visualization.

\begin{figure}
	\centering
	\includegraphics{Fig/ol_telemetries.pdf}
	\caption{}
	\label{fig:ol_telemetries}
\end{figure}

From the first panel, it can be observed that all configurations maintain a lower mean value of the longitudinal speed compared to the nominal trajectory, resulting in an increased sector time. The two configurations incorporating the robustified track limit constraint exhibit, as expected, a significantly altered CoM trajectory. This behavior explains the higher longitudinal speed observed around $\al = 0.71$, which is followed by a sharp reduction.

@@

\begin{figure}
	\centering
	\includegraphics{Fig/ol_saturation.pdf}
	\caption{}
	\label{fig:ol_saturation}
\end{figure}

\subsection{Simulations with noise realization}
This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work. For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the nominal trajectory. In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.

\begin{figure}
	\centering
	\includegraphics{Fig/olcl_traj_strings.pdf}
	\caption{}
	\label{fig:traj_strings}
\end{figure}

>>>>

\begin{figure}
	\centering
	\includegraphics{Fig/gamma_sensitivity.pdf}
	\hfill
	\includegraphics{Fig/Pzero_sensitivity.pdf}
	\caption{Effect of $\ga^\textrm{TLC}$ (left panel) and $P_0$ (right panel) on optimal trajectory. The parameter $\ga^\textrm{TLC}$ span in the range $\left[0,3\right]$ corresponding to a probability to meet the track limit constraint of 50\% and 99\%, respectively. In the left panel, in addition to the baseline value $\bar{\bP}_0$ (black line), two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ (blue line) and $\bP_0 = \frac{1}{4}\bar{\bP}_0$ (red line). These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector.}
	\label{fig:ol_sensitivities}
\end{figure}

\subsection{Robustified constraints comparison}
It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style. This analysis compares the not-robust trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC). All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of meeting both constraints.

Figure~\ref{fig:ol_telemetries} presents the variation of four signals with respect to the not-robust trajectory, for the three previously mentioned cases. The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel). Except for the $u$ panel, each plot includes dashed black lines that indicate the sign of the reference signal-whether it is positive, negative, or numerically zero. These sign indicators are represented as three distinct levels---high, zero, and low---corresponding to positive, zero, and negative values of the reference signal, respectively. The levels are scaled appropriately in each plot to ensure clarity of visualization.

\begin{figure}
	\centering
	\includegraphics{Fig/ol_telemetries.pdf}
	\caption{Variation of four signals w.r.t. the not-robust trajectory for the three configurations TLC (green line), FLC (red line) and TLC+FLC (blue line). The panels display the longitudinal speed $u$ (first panel), wheel steering angle $\de$ (second panel), total longitudinal force $F_x$ (third panel), and lateral deviation of the center of mass from the track centerline $e$ (fourth panel), over the curvilinear abscissa $\al$. The dashed lines in the last three panels indicate the sign of the reference signal, allowing one to determine whether a positive variation with respect to the baseline corresponds to an increase or a decrease in the absolute value. This distinction is particularly important for quantities that depend on the direction of the turn, such as the wheel steering angle $\de$ and the lateral deviation $e$, while in the case of $F_x$, it indicates whether the vehicle is accelerating or braking.}
	\label{fig:ol_telemetries}
\end{figure}

From the first panel, it can be observed that all configurations maintain a lower mean value of the longitudinal speed compared to the reference trajectory, resulting in an increased sector time. The two configurations incorporating the robustified track limit constraint exhibit, as expected, a significantly altered CoM trajectory. This behavior explains the higher longitudinal speed observed around $\al = 0.71$, which is followed by a sharp reduction.

@@

\begin{figure}
	\centering
	\includegraphics{Fig/ol_saturation.pdf}
	\caption{Tyre usage for front axle (first panel) and rear axle (second panel) over curvilinear abscissa $\al$. The plotted lines represent the quantity $\left[ \left( \frac{X_j(\bx,\bu)}{\mu_{x,j}} \right)^2+ \left( \frac{Y_j(\bx,\bu)}{\mu_{y,j}}\right)^2\right]\frac{1}{Z_j^2(\bx,\bu)}$, where $j=1$ denotes the front axle and $j=2$ the rear axle. The dashed black line corresponds to the reference (not-robust) configuration, while the green, red, and blue lines represent the robust configurations TLC, FLC, and TLC+FLC, respectively.}
	\label{fig:ol_saturation}
\end{figure}

\subsection{Simulations with noise realization}
This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work. For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the not-robust trajectory. In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.

\begin{figure}
	\centering
	\includegraphics{Fig/olcl_traj_strings.pdf}
	\caption{Simulations with random initial conditions and Gaussian noise realizations. In the left panel, the reference trajectory is the non-robust one, tracked using an LQR controller designed around it. In the right panel, the reference is a robust trajectory obtained using the closed-loop optimization method, and it is tracked using the corresponding optimized controller produced by the same method.}
	\label{fig:traj_strings}
\end{figure}
